Global and local DPG test functions for convection-diffusion

### 2012-10-29 docs/Jesse/Small talks/globalLocal/globalLocalTalk.tex
adding small talks folder

\title[A DPG method for Compressible Flow]{A Discontinuous Petrov-Galerkin method for compressible flow problems}
\date{August 15, 2012}
\author[J. Chan, T. Ellis, L. Demkowicz, R. Moser]{Jesse Chan, Truman Ellis \\ Supervisors: Leszek Demkowicz, Robert Moser}

@@

\frame{
\frametitle{Compressible Navier-Stokes equations}
\begin{columns}[c]
\begin{column}{.475\textwidth}
Numerical difficulties:
\begin{itemize}
\item{} Resolving solution features (sharp, localized viscous-scale phenomena)
\begin{itemize}
\item{} Shocks
\item{} Boundary layers - resolution needed for drag/load
\item{} Turbulence (non-localized)
\end{itemize}
\item{} Nonlinear convergence and uniqueness of solutions
\item{} Stability of numerical schemes
\begin{itemize}
\item{} Coarse/adaptive grids
\item{} Higher order
\end{itemize}
\end{itemize}
\end{column}
\begin{column}{.475\textwidth}
\begin{figure}
\centering
\includegraphics[scale = .7]{figs/compFlowPics/airfoil_mesh.png}
\end{figure}
\end{column}
\end{columns}
\vspace{.2cm}
\begin{center}
Idea: begin first with the model problem of convection-diffusion.
\end{center}
}
\frame{
\frametitle{Robustness: convection-diffusion as a model problem}
\vspace{-.5cm}
\[
\div \left(\beta u\right) - \epsilon \Delta u = f, \quad \text{on }\Omega \in \mathbb{R}^3
\]
\vspace{-.5cm}
\begin{columns}[c]
\begin{column}{.49\textwidth}
In 1D: $\beta u' - \epsilon u'' = f$. Standard continuous Galerkin variational formulation: solve
\[
b(u,v) = \ell(v), \quad u,v\in H^1_0([0,1])
\]
where
\begin{align*}
b(u,v) &= \int_\Omega -\beta uv' + \epsilon u'v' \\
\ell(v) &= \int_\Omega f v 
\end{align*}
\end{column}
\begin{column}{.49\textwidth}
\begin{figure}
\centering
\includegraphics[scale=.35]{figs/GalerkinOscTight.png}
\caption{Solution for $f=1$. Oscillations in the standard Galerkin method for large Peclet numbers ${\rm Pe} \coloneqq \frac{h}{\epsilon}$.}
\end{figure}
\end{column}
\end{columns}
}

\frame{
\frametitle{Streamline-upwind Petrov-Galerkin (SUPG)}
SUPG solves $b_{\rm SUPG}(u,v) = l_{\rm SUPG}(v)$, where
\begin{align*}
b_{\rm SUPG}(u,v) &= b(u,v)+ \sum_{K} \int_{K} \tau \LRp{L_{\rm adv} v} L u \\
l_{\rm SUPG}(v) &= \ell(v) + \sum_{K} \int_{K} \tau \LRp{L_{\rm adv}v} f.
\end{align*}
\vspace{-.5cm}
\begin{columns}[c]
\begin{column}{.44\textwidth}
\vspace{-.5cm}
\begin{itemize}
\item{} $Lu = \div\left(\beta u\right) - \epsilon \triangle u$, $L_{\rm adv}u = \div \left(\beta u\right)$, and $\tau$ is a parameter. 
%\item{} Recovers ``exact'' artificial diffusion for $f=0$.
\item{} Effective for $f\neq 0$. %``exact'' artificial diffusion.
\item{} \emph{Residual-based} stabilization. 
\end{itemize}
\end{column}
\begin{column}{.54\textwidth}
\begin{figure}
\centering
\includegraphics[scale=.38]{figs/SUPG.png}
%\caption{Exact, Bubnov-Galerkin, and SUPG solutions.}
\end{figure}
\end{column}
\end{columns}
}

\frame{
\begin{columns}
\begin{column}{.49\textwidth}
Can be interpreted as a \emph{Petrov-Galerkin} method,
\[
b\LRp{u,\tilde{v}_{i}} = \ell\LRp{\tilde{v}_{i}}, \quad \forall i = 1,\ldots,N-1,
\]
where the SUPG test function $\tilde{v}_{i}$ is defined elementwise as\footnotemark
\[
\tilde{v}_{i} = \phi_i(x) + \tau L_{\rm adv} \phi_i.  
\]
\end{column}
\begin{column}{.49\textwidth}
\begin{figure}
\centering
\subfigure{\includegraphics[scale=.16]{figs/SUPGtest.png}}
\caption{SUPG test function $v_i$.}
\end{figure}
\end{column}
\end{columns}
\footnotetext{\bibentry{SUPG}}
}

\section{The DPG Method}
\frame{
\frametitle{DPG: a minimum residual method via optimal testing}
Given a trial space $U$ and Hilbert test space $V$,
\[
b(u,v) = \ell(v), \quad u\in U, \quad v\in V
\]
This is equivalent to the operator equation posed in $V'$
\[
Bu = \ell
\]
if we identify $B:U\rightarrow V'$ and $\ell\in V'$ such that 
\begin{align*}
\langle Bu,v\rangle_V &\coloneqq b(u,v), \quad u\in U, v\in V,\\
\langle \ell,v\rangle_V &\coloneqq \ell(v), \quad v\in V.
\end{align*}
We seek to minimize the dual functional over $U_h \subset U$
\[
J(u_h) = \frac{1}{2}\|Bu_h-\ell\|_{V'}^2 \coloneqq\frac{1}{2} \sup_{v\in V\setminus\{0\}} \frac{| b(u_h,v)-\ell(v)|^2}{\nor{v}_V^2}.
\]
}

\begin{frame}
Let $R_V: V \to V'$ be the isometric Riesz map st.\
\[
\langle R_V v,\delta v\rangle_V \coloneqq(v, \delta v)_V, \quad \forall \delta v \in V.
\]
Then, our functional $J(u_h)$ is equal to
\begin{equation*}
\min_{u_h\in U_h} J(u_h) = \frac{1}{2}\left\|Bu_h-\ell\right\|_{V'}^2 =  \frac{1}{2}\left\|R_V^{-1}(Bu_h-\ell)\right\|_V^2.
\end{equation*}
First order optimality: G\^ateaux derivative is zero in all directions $\delta u \in U_h$
\begin{align*}
\left(R_V^{-1}(Bu_h-\ell),R_V^{-1}B\delta u\right)_V &= 0, \quad \forall \delta u \in U_h.\\
\rightarrow \LRa{(Bu_h-\ell),R_V^{-1}B\delta u} &= 0, \\
\rightarrow b(u_h,R_V^{-1}B\delta u) - \ell(R_V^{-1}B\delta u) &= 0
\end{align*}
\end{frame}

\frame{
\frametitle{Summary: select test functions to minimize residuals}
For $\delta u \in U_h$, define the \textcolor{red}{optimal test function} $v_{\delta u}$. 
\begin{equation*}
v_{\delta u} \coloneqq R_V^{-1}B\delta u.
\end{equation*} 
Then, the residual $J(u_h) = \frac{1}{2}\left\|Bu_h-\ell\right\|_{V'}^2$ is minimized by the solution of
\[
b\LRp{u_h,v_{\delta u}} = \ell\LRp{v_{\delta u}}, \quad \forall \delta u \in U_h.
\]
}

\frame{
\frametitle{Practical details of DPG}
Computation of $v_{\delta u} \coloneqq R_V^{-1}B\delta u$ is \textbf{global} and \textbf{infinite-dimensional}.
\begin{itemize}
\item By choosing a \textcolor{red}{broken} test space $V$ and \textcolor{red}{localizable} norm $\nor{v}_V^2 = \sum_K \nor{v}_{V(K)}^2$, test functions can be determined locally.
\item In practice, we use an \textcolor{red}{enriched space} $V_h \subset V$, where $\dim(V_h) > \dim(U_h)$ elementwise, and \textcolor{red}{optimal test functions} are approximated by computing $v_{\delta u} \coloneqq R_{V_h}^{-1}B\delta u \in V_h$ through\footnote{\bibentry{DPG2}}
\[
\LRp{v_{\delta u},\delta v}_V = b(\delta u,\delta v), \quad \delta u\in U_h, \quad \forall \delta v\in V_h
\]
Typically, if $U_h = \mathcal{P}^p(\mathbb{R}^n)$, $V_h = \mathcal{P}^{p+\triangle p}(\mathbb{R}^n)$, where $\triangle p \geq n$.\footnote{\bibentry{practicalDPG}} 
\end{itemize}
%\footnotetext{\bibentry{DPG2}}
}

\frame{
\frametitle{Properties of DPG}
DPG provides a symmetric positive-definite stiffness matrix. Let $\{\phi_j\}_{j=1}^m$ be a basis for $U_h$, and $\{v_i\}_{i=1}^n$ a basis for $V_h$, such that $n>m$. Then, for 
\begin{align*}
B_{ji} &= b(\phi_j,v_i),\\
l_i &= \ell(v_i),
\end{align*}
DPG solves the discrete system for degrees of freedom $u$
\begin{align*}
\LRp{B^T R_V^{-1}B} u  &= \LRp{B^T R_V^{-1}} l,
\end{align*}
where, under a localizable norm and discontinuous test functions, $R_V^{-1}$ is block diagonal.
}

\frame{
\frametitle{Properties of DPG}
Additional properties of DPG include\footnotemark
\begin{itemize}
\item DPG provides the best approximation in the \textcolor{red}{energy norm}
\[
\|u\|_E = \|Bu\|_{V'} = \sup_{\nor{v}_V=1} \left|b(u,v)\right|.
\]
\item The energy error is computable through the residual
\[
\nor{u-u_h}_E = \nor{B(u-u_h)}_{V'} = \nor{R_V^{-1}(l-Bu_h)}_V = \nor{e}_V
\]
where the \textcolor{red}{error representation function} $e$ is defined through $(e,\delta v)_V = \ell(v)-b(u_h,\delta v)$ for all $\delta v\in V$. 
\end{itemize}
\footnotetext{\bibentry{DPG3pub}}
}

%\frame{
%\frametitle{Duality of norms}
%Duality of norms: for any trial norm $\nor{u}_U$,
%\[
%\nor{u}_{U} = \sup_{v \in V}\frac{b\LRp{u,v}}{\nor{v}_{V,U}}, \quad \nor{v}_{V,U} = \sup_{u \in U}\frac%{b\LRp{u,v}}{\nor{u}_{U}}.
%\]
%Likewise, for any test norm $\nor{v}_V$, 
%\[
%\nor{v}_{V} = \sup_{u \in U}\frac{b\LRp{u,v}}{\nor{U}_{U,V}}, \quad \nor{u}_{U,V} = \sup_{v \in V}\frac%{b\LRp{u,v}}{\nor{v}_{V}}.
%\]
%
%}

\frame{
\frametitle{Ultra-weak formulation}
\vspace{-.75cm}
\begin{columns}
\begin{column}{.64\textwidth}
Given a first order system $Au = f$, we identify the \textbf{\emph{partition}} $\Oh$ and \textbf{\textcolor{red}{mesh skeleton}} \textcolor{red}{$\Gh$}.
\end{column}
\begin{column}{.34\textwidth}
%\[
%\Oh = \cup_{j=1}^\Nel K_j, \quad \Gh = \cup_{j=1}^\Nel \partial K_j
%\]
\begin{figure}[!h]
\centering
\includegraphics[scale = .155]{mesh_skel.pdf}
\end{figure}
\end{column}
\end{columns}
\vspace{-.5cm}
The ultra-weak formulation for $Au = f$ on $\Oh$ is
\[
b\left(\left(u, \widehat{u}\right),v\right) \coloneqq \sum_K\langle \widehat{u},v \rangle_{\partial K} + (u,A_h^*v)_{\Oh}= \LRp{f,v}_{\Oh}.
\]
Under proper assumptions, $\sum_K\langle \widehat{u},v \rangle_{\partial K} = \langle \widehat{u}, \jump{v} \rangle_{\Gh}$, with energy setting
\[
u\in L^2\LRp{\Oh} \equiv L^2(\Omega), \quad v\in V=D(A^*_h), \quad
\widehat{u}\in \gamma(D(A)),
\]
where $D(A_h^*)$ is the broken graph space of the formal adjoint $A_h^*$, and $\gamma(D(A))$ the trace space of the graph space of operator $A$.\footnote{\bibentry{analysisDPG}}
}

\frame{
\frametitle{The canonical ``graph'' test norm}
Recall $\nor{u}_E \coloneqq \sup_{v\in V\setminus\{0\}}\frac{b(u,v)}{\nor{v}_V}$. Under the ultra-weak formulation, the trial norm 
\[
\nor{\LRp{u,\widehat{u}}}_U \coloneqq \|u\|^2_{\L} + \|\widehat{u}\|^2
\] 
generates a test norm \emph{equivalent} to 
\[
\nor{v}_{V_{\rm opt}} \coloneqq \|A_h^*v\|_{\L}^2 +\left(\sup_{\widehat{u}} \frac{\LRa{ \widehat{u},\jump{v} }_{\Gh}}{\|\widehat{u}\|}\right)^2.
\]
This norm is not localizable, so we instead substitute the jump terms for an $L^2$ term, giving us the \emph{graph norm}\footnote{\bibentry{Bui-ThanhDemkowiczGhattas11b}}
\[
\nor{v}_V \coloneqq \|A_h^*v\|_{\L}^2 + \nor{v}_{L^2}.
\]
}

\frame{
\frametitle{Goal: robust higher order adaptive methods}
\begin{columns}
\begin{column}{.45\textwidth}
Our aim is to design an adaptive method for compressible laminar flows that is 
\begin{itemize}
\item robust in Reynolds number
\item stable for arbitrary elements
\end{itemize}

This proposal will outline
\begin{enumerate}
\item a robust DPG method for convection-diffusion,
\item extension to nonlinear problems and systems,
\item proposed work.
\end{enumerate}

\end{column}
\begin{column}{.54\textwidth}
\begin{figure}
\centering
\subfigure{\includegraphics[scale=.23]{figs/Re1000p2/u2_full.png}}
\subfigure{\includegraphics[scale=.37]{figs/Re1000p2/mesh6.png}}
\caption{$u_2$ and $p=2$ mesh for $\Reyn = 1000$.}
\end{figure}
\end{column}
\end{columns}
}

\section{A Robust DPG Method for Convection-Diffusion}

\frame{
\frametitle{Ultra-weak formulation for convection-diffusion}
In first order form, the convection-diffusion equation is
\[
A \LRp{u,\sigma} \coloneqq 
\LRs {
\begin{array}{c}
\div (\beta u - \sigma) \\ \frac{1}{\epsilon}\sigma - \grad u
\end{array}} = \LRs{
\begin{array}{c}
f \\ 0
\end{array}
}.
\]
The variational formulation is
\begin{align*}
b\left(\left(u,\sigma, \widehat{u}, \widehat{f}_n\right),
\left( v, \tau \right)\right) = \left(u,\grad_h\cdot \tau - \beta \cdot \grad_h
v\right)_{\Oh} + \left(\sigma, \epsilon^{-1} \tau + \grad_h v\right)_{\Oh}&\\
  - \LRa{\jump{\tau\cdot n}, \widehat{u} }_{\Gh} + \LRa{\widehat{f}_n,\jump{v} }_{\Gh}&,
\end{align*}
where $\widehat{f}_n \coloneqq \beta_n u - \sigma_n$ and $\LRa{\widehat{f}_n,\jump{v} }_{\Gh}$ is defined 
\[
\LRa{\widehat{f}_n,\jump{v} }_{\Gh}\coloneqq \sum_K\int_{\partial K}{\rm sgn}(\vec{n}) \, \widehat{f}_n v.
\]
}

\frame{
\frametitle{Graph norm under convection-diffusion}
For convection-diffusion, the graph test norm is defined elementwise
\[
\|\left(v,\tau\right)\|_{V(K)}^2 = \| \div \tau - \beta \cdot \grad v
\|_{L^2(K)}^2 + \| \epsilon^{-1} \tau + \grad v \|_{L^2(K)}^2 +
\|v\|_{L^2(K)}^2.
\]
Problem with this test norm: approximability of test functions.
\vspace{-.3cm}
\begin{figure}[!h]
\centering
\includegraphics[scale=.175]{figs/opt.png}
\caption{$v$ and $\tau$ components of the 1D optimal test functions for flux $\widehat{f}_n$ on the \emph{right-hand} side of a unit element for $\epsilon = 0.01$. }
\label{fig:optTestBoundary}
\end{figure}
}

\frame{
\frametitle{Determining an alternative test norm}
Recall the convection-diffusion bilinear form
\begin{align*}
b\left(\left(u,\sigma, \widehat{u}, \widehat{f}_n\right),
\left( v, \tau \right)\right) = \left(u,\div \tau - \beta \cdot \grad
v\right)_{\Oh} + \left(\sigma, \epsilon^{-1} \tau + \grad v\right)_{\Oh}&\\
 - \LRa{\jump{\tau\cdot n}, \widehat{u} }_{\Gh} + \LRa{\widehat{f}_n,\jump{v} }_{\Gh}&,
\end{align*}
We recover $\nor{u}^2_{L^2(\Omega)}$ by choosing conforming $\LRp{v,\tau}$ satisfying the \emph{adjoint equations}
\begin{align*}
\div\tau - \beta \cdot \grad v   &= u\\
\frac{1}{\epsilon}\tau + \grad v &= 0
\end{align*}
with boundary conditions s.t.\ $\LRa{\jump{\tau\cdot n}, \widehat{u} }_{\Gamma}$ and $\LRa{ \widehat{f}_n,\jump{v} }_{\Gamma}$ vanish.
}

\frame{
\frametitle{A robust bound}
``Necessary'' conditions for robustness --- let $\boldsymbol U = \LRp{u,\sigma,\widehat{u},\widehat{f}_n}$. Then, by choosing specific $\LRp{v,\tau}$ satisfying the adjoint equations,
\[
\nor{u}^2_{L^2(\Omega)} = b\LRp{{\boldsymbol U},\LRp{v,\tau}} = \frac{b\LRp{{\boldsymbol U},\LRp{v,\tau}}}{\nor{\LRp{v,\tau}}_V} \nor{\LRp{v,\tau}}_V \leq \nor{\boldsymbol U}_E \nor{\LRp{v,\tau}}_V
\]
Let $\lesssim$ denote a robust bound - \textcolor{red}{if $ \nor{\LRp{v,\tau}}_V \lesssim \|u\|_{L^2(\Omega)}$}, then we have that
\[
\nor{u}_{L^2(\Omega)} \lesssim \nor{\boldsymbol U}_E
\]
\textbf{Main idea: the \textcolor{red}{test norm} should measure \textcolor{red}{adjoint solutions} robustly.}
}

\frame{
\frametitle{Choice of inflow boundary condition}

\begin{columns}
\begin{column}{.49\textwidth}
We impose the standard outflow wall boundary condition on $u$. For inflow boundary condition: 
\begin{itemize}
\item{} The standard choice of inflow boundary condition: $u = u_0$.
\item{} We impose the non-standard inflow condition: $\widehat{f}_n \coloneqq \beta_n u - \sigma_n \approx \beta_n u_0$ on $\Gamma_{\rm in}$.
\end{itemize}
\end{column}

\begin{column}{.49\textwidth}
\begin{figure}
\includegraphics[scale=.23]{figs/primal.png}
\caption{Non-standard inflow.}
\end{figure}
\end{column}
\end{columns}
When $\sigma_n \approx 0$ near the inflow, condition on $\widehat{f}_n$ approximates condition on $u$.

@@

\frame{
\frametitle{Dirichlet inflow boundary condition}
Standard choice of boundary condition: $u = u_0$ on inflow boundary $\Gamma_{\rm in}$, induces boundary layers in adjoint problems, \textcolor{red}{$\nor{\beta\cdot \grad v}_{L^2} = O(\epsilon^{-1})$}.
\begin{figure}[h!]
\centering
\subfigure[Primal problem]{\includegraphics[scale=.23]{figs/primalDir.png}}
\subfigure[Adjoint problem]{\includegraphics[scale=.23]{figs/adjointDir.png}}
\caption{For the standard Dirichlet inflow condition, the solution to the adjoint problem can develop strong boundary layers at the outflow of the adjoint problem. }
\end{figure}
}

\frame{
\frametitle{New inflow boundary condition on $\widehat{f}_n$}
Non-standard choice of boundary condition: $\widehat{f}_n = \beta_n u_0$ on $\Gamma_{\rm in}$, induces smoother adjoint problems, \textcolor{red}{$\nor{\beta\cdot \grad v}_{L^2} = O(1)$}.
\begin{figure}[h!]
\centering
\subfigure[Primal problem]{\includegraphics[scale=.23]{figs/primal.png}}
\subfigure[Adjoint problem]{\includegraphics[scale=.23]{figs/adjoint.png}}
\caption{Under the new inflow condition, the wall-stop boundary condition is relaxed to a zero-stress condition at the outflow boundary of the adjoint problem.}
\end{figure}
}

\frame{
\frametitle{Test norms and adjoint solutions}
\textbf{Intuition:} the effectiveness of DPG under a test norm is governed by how a \textbf{specific test norm} measures the \textbf{solutions of the adjoint problem}.

\begin{figure}[!h]
\centering
\subfigure[Dirichlet inflow]{\includegraphics[scale=.35]{figs/dirichletBC.png}}
\subfigure[``Convection'' inflow]{\includegraphics[scale=.35]{figs/newBC.png}}
\caption{DPG solutions to convection-diffusion for both inflow conditions using an $H^1$ test norm.}
\end{figure}
}

%\frame{
%\frametitle{``Building blocks'' norm}
%Test norm quantities are robustly bounded from above by $\nor{u}_{L^2(\Omega)}$.  
%\[
%\|\left(v,\tau\right)\|_{V,K}^2 = \|v\|^2 + \epsilon \|\grad v\|^2 + \|\beta \cdot \grad v\|^2 + \| \div \tau\|^2 + \frac{1}{\epsilon}\|\tau\|^2.
%\]
%Problem: boundary layers in optimal test functions:
%\begin{figure}[!h]
%\centering
%\subfigure{\includegraphics[scale=.15]{figs/mesh1.png}}
%\subfigure{\includegraphics[scale=.15]{figs/sol1.png}}
%\caption{The $v$ component of the optimal test function corresponding to flux $\widehat{u} = x(1-x)$ on the bottom side of a unit element for $\epsilon = 0.01$.}
%\label{fig:boundaryTest}
%\end{figure}
%}
\frame{
\frametitle{Adjoint estimates, or ``norm building blocks''}
For solutions $(v,\tau)$ of the adjoint equations, the following quantities are robustly bounded from above by $\nor{u}_{L^2(\Omega)}$.  
\begin{align*}
&\|v\|, \sqrt{\epsilon} \|\grad v\|, \|\beta \cdot \grad v\| \\
&\| \div \tau\|, \frac{1}{\sqrt{\epsilon}}\|\tau\|.
\end{align*}
We will construct a test norm through a combination of the above quantities, such that
\begin{itemize}
\item $v$ and $\tau$ decoupled (no systems).
\item Coefficients are of equal order after transforming to the unit element (no boundary layers).
\end{itemize}
}
\frame{
\frametitle{Mesh-scaled test norms}
Our test norm, as defined over a single element $K$, is now
\begin{align*}
\|\left(v,\tau\right)\|_{V,K}^2 = \min\left\{\frac{\epsilon}{|K|},1\right\}\|v\|^2 + \epsilon \|\grad v\|^2 + \|\beta \cdot \grad v\|^2 +&\\
\| \div \tau\|^2 + \min\left\{\frac{1}{\epsilon},\frac{1}{|K|}\right\}\|\tau\|^2&.
\end{align*}
which induces the proven \emph{robust} bound\footnote{\bibentry{DPGrobustness2}}
\[
\nor{u}_{L^2(\Omega)}+ \nor{\sigma}_{L^2(\Omega)} + \epsilon\nor{\widehat{u}} + \sqrt{\epsilon} \nor{\widehat{f}_n} \lesssim \nor{\LRp{u,\sigma,\widehat{u},\widehat{f}_n}}_E.
\]
}

\frame{
\frametitle{Erikkson-Johnson model problem}

On domain $\Omega = [0,1]^2$, with $\beta = (1,0)^T$, $ f = 0$ and boundary conditions
\begin{align*}
\widehat{\beta_nu - \sigma_n} &= \widehat{f}_n = u_0, \quad \beta_n \leq 0\\
\widehat{u} &= 0, \quad \beta_n > 0
\end{align*}

%Separation of variables gives an analytic solution. 
All numerical experiments done using Camellia codebase.\footnote{\bibentry{Camellia}}
%\[
%u(x,y) = C_0 + \sum_{n=1}^\infty C_n \frac{\exp(r_2(x-1)-\exp(r_1(x-1)))}{r_1\exp(-r2) - r_2\exp(-r1)}\cos(n\pi y)
%\]
\begin{figure}
\centering
\subfigure{
\includegraphics[scale=.23]{figs/wallBC_exact_u.png}
}
\subfigure{
\includegraphics[scale=.23]{figs/wallBC_exact_sigx.png}
}
\subfigure{
\includegraphics[scale=.23]{figs/wallBC_exact_sigy.png}
}
\caption{Exact solution for $u$, $\sigma_x$, and $\sigma_y$ for $\epsilon = .01$, $C_1 = 1$, $C_n=0$, $n\neq 1$}
\end{figure}
}

\frame{
\frametitle{Error rates}
\begin{figure}
\centering
\subfigure{
\includegraphics[scale=.33]{figs/errorrates_wallBC.png}
}
\subfigure{
\includegraphics[scale=.33]{figs/L2energyratio_wallBC.png}
}
%\subfigure{
%\includegraphics[scale=.22]{figs/u_pointdiff_wallBC.png}
%}
\caption{$L^2$/energy errors and their ratio for $\epsilon = .01, .001, .0001$.}
\end{figure}
}

%\frame{
%\frametitle{Regularization}
%Discontinuous inflow conditions on $\widehat{f_n}$, out of alignment with mesh. Boundary condition $\epsilon \grad u \cdot n = 0$ at outflow.
%\begin{figure}
%\centering
%\subfigure{
%\includegraphics[scale=.33]{figs/discontinuous.png}
%}
%\subfigure{
%\includegraphics[scale=.305]{figs/discontinuousL2rates.png}
%}
%\caption{Discontinuous hat inflow data with $\epsilon = 1e-5$, convection solution errors.}
%\end{figure}
%}
\frame{
\frametitle{Local conservation}
Given stiffness matrix $K$, can introduce Lagrange multipliers to enforce element-wise constraints, leading to the saddle point problem
\[
\arr{K}{B}{B^T}{0}\vecttwo{u}{\lambda} = \vecttwo{f}{c}.
\]
where the Lagrange multipliers $\lambda$ can be locally condensed. We use the test norm
\begin{align*}
\|\left(v,\tau\right)\|_{V,K}^2 = \left|\int_K v\right|^2 + \epsilon \|\grad v\|^2 + \|\beta \cdot \grad v\|^2 +&\\
\| \div \tau\|^2 + \min\left\{\frac{1}{\epsilon},\frac{1}{|K|}\right\}\|\tau\|^2&.
\end{align*}
}

\frame{
\frametitle{Additional examples}
\begin{figure}
\centering
\subfigure[Advection skew]{
\includegraphics[scale = .175]{figs/Conservative/hughes.png}
}
\subfigure[Flow over a step]{
\includegraphics[scale = .175]{figs/Conservative/step.png}
}
\subfigure[Double glazing]{
\includegraphics[scale = .182]{figs/Conservative/doubleGlazing.png}
}
\end{figure}
Locally conservative version of DPG performs very similarly to the standard DPG in all cases over a range of $\epsilon$. 
}

\frame{
\frametitle{Regularization}
For $\beta = (-y,x)^T$ on $\Omega = [-1,1]^2$. Ill posed in the convection setting. Similar tests have been done with discontinuous data.

\begin{figure}
\centering
\subfigure{
\includegraphics[scale=.21]{figs/vortex.png}
}
\subfigure{
\includegraphics[scale=.2015]{figs/vortexMesh.png}
}
\caption{Steady vortex problem with $\epsilon = 1e-4$.}
\end{figure}
\vspace{-.25cm}
}

\frame{
\frametitle{Current work - Hemker problem}

>>>>

\title[Global/local DPG]{Global and local DPG test functions for convection-diffusion}

@@

\frame{
\frametitle{Ultra-weak formulation}

Given a first order system $Au = f$, multiply by test function $v$ and integrate 
\[
\LRp{Au,v} = \LRa{\gamma\LRp{Au},v} + \LRp{u,A^*_hv} = \LRp{f,v}
\]
We identify boundary terms $\LRa{\gamma\LRp{Au},v}_{\Gh} = \LRa{\widehat{u},v}_{\Gh}$ as unknowns $\widehat{u}$ on $\Gh$. This gives us the bilinear form.
\begin{align*}
b(\LRp{u,\uh},v) &\coloneqq  \LRa{\uh,v} + \LRp{u,A^*_hv}  \\
l(v) &\coloneqq \LRp{f,v}
\end{align*}
%Let boundary conditions be applied on $\uh$ on $\Gamma_{\rm BC}$.  
DPG approximates optimal test functions  $v_{\delta u}$ for all $\delta u\in U_h$ by solving on a local level
\[
\LRp{v_{\delta u},\delta v} = b(\delta u, \delta v), \quad \delta v\in V_h(K)
\]

@@

\frame{
\frametitle{$L^2$ best approximations under the ultra-weak formulation}
If our optimal test functions satisfy for all $\delta u_h \in U_h$ 
\begin{align*}
A^*v &= \delta u_h, \quad \text{on } \Omega
%\gamma\LRp{A^*v} &= 0, \quad \text{on } \Gamma \setminus \Gamma_{\rm BC}  \\
\end{align*}
with boundary conditions on $v$ such that the boundary terms disappear, we get back the best $L^2$ approximation by virtue of 
\begin{align*}
b(\LRp{u_h,\uh_h},v) = \LRa{\uh_h,v} + \LRp{u_h,A^*v} &= \LRp{u_h,\delta u_h}\\
\LRp{f,v} = b(\LRp{u,\uh},v) =  \LRa{\uh,v} + \LRp{u,A^*v} &= \LRp{u,\delta u_h}
\end{align*}
Corresponds to a graph norm choice of test norm: under assumptions of boundedness below of $B$, for $\delta > 0$, we can define as a DPG test norm
\[
\nor{v}_V \coloneqq \nor{A^*v}_{L^2} + \delta \nor{v}_{L^2}
\]
which, as $\delta \rightarrow 0$, gives an equivalent result.  
}

\frame{
\frametitle{Globally optimal test functions}
Recall that DPG optimal test functions are from a local inversion of the Riesz operator.  We can choose a \textcolor{red}{conforming} test space and invert the Riesz operator over the entire domain: 
\begin{align*}
V_{\rm global} &\coloneqq \{v\in \oplus V_K: \LRa{\uh_h, \jump{v}}_{\Gh^0}, \forall \uh_h \in \widehat{U}_h\} \\
\LRp{v,\delta v}_{\Omega} &= b(u_h,\delta v), \quad v\in V_{\rm global}
\end{align*}
We refer to these as \textit{globally} optimal test functions.\footnote{Selecting these test functions removes internal traces from the big picture!}  The test space resulting from the local inversion of the Riesz operator is related to the \textit{globally} optimal test space through the following lemma: 
\begin{lemma}
The globally optimal test space is contained in the locally optimal test space.  
\end{lemma}
}

\frame{
\frametitle{Optimal test functions for convection-diffusion}
For convection-diffusion,
\begin{align*}
b\left(\left(u,\sigma, \widehat{u}, \widehat{f}_n\right),
\left( v, \tau \right)\right) = &\left(u,\grad_h\cdot \tau - \beta \cdot \grad_h
v\right)_{\Oh} + \left(\sigma, \epsilon^{-1} \tau + \grad_h v\right)_{\Oh}\\
  &- \LRa{\jump{\tau\cdot n}, \widehat{u} }_{\Gh} + \LRa{\widehat{f}_n,\jump{v} }_{\Gh},
\end{align*}
where 
\begin{align*}
\widehat{f}_n \coloneqq \beta_n u - \sigma_n \in H^{-1/2}(\Gh), \quad \widehat{u} \in H^{1/2}(\Gh)
\end{align*}
The adjoint problem for $L^2$ best optimality under convection diffusion is
\begin{align*}
\left.
\begin{array}{cl}
\div\tau - \beta \cdot \grad v   &= u\\
\frac{1}{\epsilon}\tau + \grad v &= 0
\end{array}
\right. ,
\end{align*}
with boundary condition $v = 0$ on $\Gamma$. \textcolor{red}{A boundary layer forms at the inflow.}
}

\frame{
\frametitle{The graph norm for convection-diffusion}
The graph norm for convection diffusion forms boundary layers on each element, which are only resolvable using special subgrid meshes.  However, 
\begin{enumerate}
\item The globally optimal test functions only have boundary layers at the boundary
\item The test functions for $L^2$ optimality only have boundary layers at the \textcolor{red}{inflow} boundary
\end{enumerate}
Question: do we need to resolve the boundary layers in optimal test functions everywhere?
}
\frame{
\frametitle{Numerical experiment}
\begin{itemize}
\item Given the Erikkson-Johnson problem, we use the graph test norm and compute both energy and $L^2$ errors.  
\item We then refine near the inflow and compare the energy and $L^2$ errors.  
\end{itemize} 
Best approximation error is small near the inflow, so changes in energy/$L^2$ error are due to resolution of test functions, not the solution.  
\begin{figure}
\centering
\subfigure{\includegraphics[scale = .35]{figs/qoptInflowDirichlet.png}}
\subfigure{\includegraphics[scale=.35]{figs/qoptInflowDirichletRatio.png}}
%\caption{Energy error and total $L^2$ error and their ratios for different levels of resolution of the inflow layer.}
\end{figure}
}

\frame{
\frametitle{Distribution of error}
We expect that the term $\nor{\div \tau - \beta \cdot \grad v}_{L^2}$ is bounded uniformly in $\epsilon$. However, the term $\nor{\frac{1}{\epsilon}\tau - \grad v}_{L^2}$ is not.\footnote{Analytical calculations in 1D show that this term grows with the Peclet number in the element at the inflow, where the adjoint solution develops a boundary layer.}

As the boundedness of this term determines the robustness of $\sigma$, our energy norm is
\begin{align*}
\nor{\boldsymbol U}_E &\coloneqq \sup_{\LRp{v,\tau}}\LRp{\frac{\LRp{u,\Grad_h \cdot \tau - \beta \cdot \Grad_h v}_{\Omega}}{\nor{\LRp{v,\tau}}_V} + \frac{\LRp{\sigma, \frac{1}{\epsilon}\tau + \Grad_h v}_{\Omega}}{\nor{\LRp{v,\tau}}_V}} + \text{etc} \\
&\approx \nor{u}_{L^2} + C_{\text{Pe}}\nor{\sigma}_{L^2} + \nor{\LRp{\uh,\fnh}}
\end{align*}
where $\text{Pe}$ is the element Peclet number near the inflow.  
%For the quasi-optimal test norm, underresolution of the inflow boundary layer negatively affects only the robustness of $u$. 
}

\frame{
\frametitle{Three levels of test functions}
\begin{itemize}
\item $L^2$ optimal test functions resulting from a global adjoint problem,
\item Global test functions resulting from a global Riesz inversion,
\item Local test functions resulting from local Riesz inversions.
\end{itemize}

Thoughts:
\begin{itemize}
\item The complete local resolution of test functions may not be necessary
\item You can have optimal test functions with boundary layers even when there are none in the global adjoint: resolution of these on a global level may still be necessary to maintain robustness.
\end{itemize}
